---
title: 'Final Project: Predicting World CUp Winner'
author: "Rodrigue Beleho"
date: "December 20, 2018"
output: ioslides_presentation
---


##We decided to use the FIFA World Cup Matches dataset in order to predict the winner of the 2018 World Cup.

```{r,include=FALSE}
Iaquinta = read.csv("C:\\Users\\student\\Desktop\\MATH 421\\Math 421 Final Project\\WorldCupMatches.csv")
library(ggplot2)
library(caret)
library(rpart)
library(rattle)
library(lattice)
```


##Removing useless columns
```{r}
Iaquinta[,"RoundID"] = NULL
Iaquinta[,"MatchID"] = NULL
Iaquinta[,"Referee"] = NULL
Iaquinta[,"Assistant.1"] = NULL
Iaquinta[,"Assistant.2"] = NULL
Iaquinta[,"Datetime"] = NULL
Iaquinta[,"Home.Team.Initials"] = NULL
Iaquinta[,"Away.Team.Initials"] = NULL
```


##CHecking for missing values
```{r}
sum(is.na(Iaquinta))
```


##DISCUSSION ON MISSING VALUES


##Handling missing values (1) - Missing values of categorical variables are replaced by the most frequent category in the variables
```{r}
AL=function(x){
  for (i in 1:ncol(x)){
    if (is.numeric(x[,i])){
      x[,i][is.na(x[,i])]=mean(x[,i], na.rm=TRUE)
    }else{
      levels=unique(x[,i])
      x[,i][is.na(x[,i])]=levels[which.max(tabulate(match(x[,i], levels)))]
    }
  }
  return (x)
}
Iaquinta <- AL(Iaquinta)


#We had 22322 missing values in the first place
#This method brings the #of missing values to 0
```


##Taking Care of the levels
```{r}
levels(Iaquinta$Stage)
levels(Iaquinta$Stage)=c("Prelim", "Final", "Prelim", "Prelim","Prelim","Prelim","Prelim","Prelim","Prelim","Prelim","Prelim","Prelim","Prelim","Prelim","Prelim","Prelim","Prelim","Semi_Final","Semi_Final","Prelim","Quarter_Final","Round_16","Semi_Final","Semi_Final")
levels(Iaquinta$Stage)
```


##Recoding categorical variable using one hot encoding (dummy encoding)
```{r}
dummies_model <- dummyVars(Year ~., data=Iaquinta)
trainData_mat <- predict(dummies_model, newdata =Iaquinta)

trainData <- data.frame(trainData_mat)
trainData$Year <- Iaquinta$Year
```


##VISUALIZATION AND GRAPHS
```{r}
library(ggplot2)
ggplot(data = Iaquinta) + geom_density(mapping = aes(x = Attendance, fill = Stage)) + facet_wrap(~Stage)
```

```{r}
library(ggplot2)
ggplot(data = Iaquinta) + geom_density(mapping = aes(x = Home.Team.Goals, fill = Stage)) + facet_wrap(~Stage)
```



```{r}
library(ggplot2)
ggplot(data = Iaquinta) + geom_density(mapping = aes(x = Away.Team.Goals, fill = Stage)) + facet_wrap(~Stage)
```


```{r}
library(ggplot2)
ggplot(data = Iaquinta) + geom_density(mapping = aes(x = Attendance, fill = Year)) + facet_wrap(~Year)
```

```{r}
library(ggplot2)
ggplot(data = Iaquinta) + geom_density(mapping = aes(x = Half.time.Home.Goals, fill = Stage)) + facet_wrap(~Stage)
```

```{r}
library(ggplot2)
ggplot(data = Iaquinta) + geom_density(mapping = aes(x = Half.time.Away.Goals, fill = Stage)) + facet_wrap(~Stage)
```


```{r}
Al22 <- function(Iaquinta,var1,var2) {
  rt = ggplot(data=Iaquinta) + geom_bar(mapping = aes(x = Iaquinta[,var1], fill = Iaquinta[,var2]), position = "dodge")
  return(rt)
}

Al22(Iaquinta, 2, 2)

```

```{r}
Al23 <- function(Iaquinta,var1,var2) {
  rt = ggplot(data=Iaquinta) + geom_density(mapping = aes(x = Iaquinta[,var1], fill = Iaquinta[,var2]), position = "dodge")
  return(rt)
}

Al23(Iaquinta, 2, 2)
```


```{r}
Al25=function(x){
  for (i in 1:ncol(x)){
    if (is.numeric(x[,i])){
      print(ggplot(data=x)+geom_density(mapping=aes(x=x[,i]))+xlab(names(x)[i]))
    }
  }
}
Al25(Iaquinta)

```

##Model Tuning - Random Forest with 10-fold cross validation
```{r}
myGrid = expand.grid(mtry = c(1:2), splitrule = c("gini"),
                     min.node.size = c(1:2))

rf_Iaquinta10 <- train(Stage~.,data = Iaquinta, method = "ranger", 
               trControl = trainControl(method ="cv", number = 10, verboseIter = TRUE),
               tuneGrid = myGrid)
rf_Iaquinta10

#Best model with an accuracy of 0.9534138 
```